### 2021-05-25 article/article.tex
update

The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux for flavor $\beta$ at detector level as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ on both probability
and flux level, we plot the flux ratios resulting from Eq.~\ref{eq:propFlux} in Fig.~\ref{fig:flux_ratio}. In the left panel, we have marked the region in which 99\% of the 
DeepCore cascade events originating from $\ne$ and $\nt$ fluxes are contained. In the right panel, we show the two regions in which 99\% of the IceCube 
and DeepCore track events originating from $\nm$ fluxes are contained. Starting with the $\nm$ flux ratio, we see that the only clear signal discernible to the IceCube detector
is a energy-independent flux deficiency with a factor in the order of $\sim 10^{0.5}$ from core-crossing neutrinos within a zenith range of $\ztrue < -0.87$. DeepCore on the other hand, 
is exposed to multiple fringes of flux surpluses with a factor in the order of $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensitive IceCube deficiency.
%TODO: check values above after final plot of flux ratios
In IceCube, we will only consider muon track events. Thus, we have to resort to the DeepCore/PINGU detector alone for the $\nu_e$ and $\nu_\tau$ fluxes.
Here we see a somewhat weaker signal, this time a zenith-independent deficiency. Where we earlier had surpluses with a factor in the order of $\sim 10$, we now have deficiencies in the order of $\sim 10^{0.5}$.

@@

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.8\linewidth]{figures/flux_ratio.pdf}
   \end{center}
   \caption{Ratio of detector NSI to SI atmospheric fluxes. We set $\emt = -0.05$, and all other $\epsilon_{\alpha\beta}=0$. 
   Within the dotted (dashed) boundaries, 99\% of the  DeepCore (IceCube) MC events are contained. Since we limit our IceCube study to muon tracks,
   a dashed region is omitted in the left panel.}\label{fig:flux_ratio}
\end{figure}%TODO: make prettier cmap

@@

The data include 14901 track-like events and 26001 cascade-like events, both divided into eight 
$ \log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins. Each event has a Monte Carlo weight $w_{ijk,\beta}$
from which we can construct the event count as
\begin{align}\label{eq:MCevents}
    N_{ijk} &= C_{ijk}\sum_{\beta}w_{ijk,\beta}\, \phi_\beta^\text{det}\,,
\end{align}
where $C_{k\beta}$ is the correction factor from the detector systematic uncertainty and $\phi_\beta^\text{det}$ is defined as Eq.~\ref{eq:propFlux}. We have now substituted the effect of the Gaussian smearing 
by treating the reconstructed and true quantities as a migration matrix.

@@

We treat the uncorrelated systematic uncertanties differently for each detector. For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
We consider best, normal, and worst-case scenarios in IceCube using
$f=5\%$, $7.5\%$, and $f=15\%$ respectively. For PINGU, we use the same form but instead use $f=0\%$, $3\%$, and $f=5\%$. %TODO: maybe source/justification for these values?
For DeepCore, we use the provided systematic error distribution which accounts for uncertainties in the finite MC statistics and the data-driven 
muon background estimate~\cite{DC2019data}. This is summarized in Table~\ref{table:syst_errors}.  %TODO segue
{\renewcommand{\arraystretch}{1.2}
\begin{table}
   \begin{tabular}{lccc}
      \hline \hline
      Experiment & Best case & Baseline & Worst case \\
      \hline
      IceCube & $5\%$ & $10\%$ & $15\%$ \\
      PINGU & $0\%$ & $3\%$ & $5\%$ \\
      \hline \hline
   \end{tabular}
   \caption{Definition of the best, baseline, and worst case scenarios considered in each experiment, modelled by $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$ with $f$ from the table.
   We do not consider different DeepCore scenarios because her systematic error distribution is already provided in the data release~\cite{DC2019data}.}\label{table:syst_errors}
\end{table}

\subsection{Constraining the NSI parameters}
First, we set all standard oscillation parameters to their current best-fit values of Eq.~\ref{eq:PINGUparams}, except for $\dm$ and $\theta_{23}$, 
which we marginalize over their $3\sigma$ ranges of \SI{2.435e-3} to \SI{2.598e-3}{\electronvolt^2} and \SI{40.1} to \SI{51.7}{\degree} respectively. %TODO: make unit ranges nicer.
After the oscillation parameters have been marginalized out, we plot $\Delta \chi^2$ for each of the four NSI parameters in Fig.~\ref{fig:joint_3D_NO}. We record the 
confidence levels in Table~\ref{table:IC_DC_results} and Table~\ref{table:PINGU_joint_results}, and best-fit points in Table~\ref{table:bestfit}.

@@

We see asymmetrical shapes in the $\Delta \chi^2$ plots in Fig.~\ref{fig:3D_NO}, especially in $\emt$ and $\eet$.%TODO: make better
We investigate this for $\emt$ by plotting the muon survival probability difference $P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$, where 
%TODO: check that this is what is plotted.
$\epsilon^\pm_{\mu\tau} = \pm 0.05$ in Fig.~\ref{fig:Pmm_asymmetry}. Muon events are the most abundant, and it suffices to study $P_{\mu\mu}$ to explain the $\emt$ asymmetry. 
We see see that positive $\emt$ generates a slightly higher $P_{\mu\mu}$ for energies around \SI{20}{\GeV} (blue area),
while negative $\emt$ produces higher $P_{\mu\mu}$ for almost all other combinations of ${\Etrue,\ztrue}$. As we see in Fig.~\ref{fig:flux_ratio}, this muon survival abundance 
is indeed preserved at flux level. So the flux for $\emt>0$ is higher than the flux for $\emt<0$. Is this still true at event level, i.e. after reconstruction? 
As we see in Fig.~\ref{fig:emt_events}, the binned PINGU event counts display strong differences for many bins, giving a high $\Delta \chi^2$ on both sides of $\emt=0$ and slightly higher for $\emt=-0.01$. DeepCore on the other hand, 
has fewer bins where the event count for $\emt=-0.01$ surpasses the event count for $\emt=+0.01$, giving weaker statistics for the negative side. Thus, we conclude that the DeepCore
$\emt$ asymmetry stems from the lower statistics for $\emt< 0 $ at probability level, which is then preserved through the flux and finally the reconstruction. %TODO: fill more conclusion?

\begin{figure}
   \begin{subfigure}{0.25\textwidth}
      \includegraphics[width=1\linewidth]{figures/Pmm_asymmetry.pdf}%TODO: redo for emt=0.01
      \caption{$P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$ using  $\epsilon^\pm_{\mu\tau} = \pm 0.05$ in the DeepCore/PINGU energy range.}\label{fig:Pmm_asymmetry}
   \end{subfigure}
   \begin{subfigure}{0.6\textwidth}
      \includegraphics[width=1\linewidth]{figures/emt_events.pdf}
   \caption{The best-fit event count ratio from each detector for $\emt = \pm 0.01$. IceCube only observes a small surplus of events for $\emt=-0.01$ compared to 
   DeepCore and PINGU due to the weak NSI effect at high \si{\GeV} energies.}\label{fig:emt_events}%TODO: improve caption?
   \end{subfigure}
   \caption{Probability difference and best-fit event count. Negative $\emt$ has the strongest signal, as the thin blue 
   region stemming from $\emt = 0.05$ is less prominent than the red regions from $\emt=-0.05$.}
\end{figure}

@@

For the joint analysis, we follow the parameter goodness-of-fit prescription~\cite{maltoni2003} and construct the joint $\chi^2$ as 
\begin{align}\label{eq:joint_chisq}
    \chi^2_\text{joint} = \sum_\text{exp}\chi^2_\text{exp} - \chi^2_\text{exp,min}\,
\end{align}
with test statistic $\chi^2_\text{joint,min}$. The $\Delta \chi^2_\text{joint}$ is then $\Delta \chi^2_\text{joint} = \chi^2_\text{joint} - \chi^2_\text{joint,min}$.
The results are shown in Fig.~\ref{fig:joint_3D} and summarized in Tables~\ref{table:IC_DC_results} and~\ref{table:PINGU_joint_results}.
Comparing the PINGU and the DeepCore results in Fig.~\ref{fig:3D_NO}, we note that the best-fit for each NSI parameter for the PINGU experiment is expected to be zero. This is because the `data' we generated during 
the PINGU simulations assume no NSI since they have yet to be observed in nature. This introduces a non-NSI bias in all joint analyses which include PINGU,
since PINGU has stronger statistics than DeepCore and will thus pull the joint $\chi^2$ towards $\epsilon =0$.
\newpage

>>>>

The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux for flavor $\beta$ at detector level as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ on both probability
and flux level, we plot the flux ratios resulting from Eq.~\ref{eq:propFlux} in Fig.~\ref{fig:flux_ratio}. In the left panel, we plot the flux in region in which 99\% of the 
DeepCore track events originating from $\nm$ fluxes are contained. In the right panel, we show the same but with IceCube events. We see that the only clear signal discernible to the IceCube detector
is a energy-independent flux deficiency with a factor in the order of $\sim 10^{0.5}$ from core-crossing neutrinos within a zenith range of $\ztrue < -0.87$. DeepCore on the other hand, 
is exposed to multiple fringes of flux surpluses with a factor in the order of $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensitive IceCube deficiency.
%TODO: check values above after final plot of flux ratios

@@

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.8\linewidth]{figures/flux_ratio.pdf}
   \end{center}
   \caption{Ratio of detector NSI to SI atmospheric $\nu_\mu + \bar{\nu}_\mu$ fluxes expressed as $(\phi_{\nu_\mu}^{NSI} + \phi_{\bar\nu_\mu}^{NSI})/(\phi_{\nu_\mu}^{SI} + \phi_{\bar\nu_\mu}^{SI})$.
   We set $\emt = -0.05$, and all other $\epsilon_{\alpha\beta}=0$. 
   Left (right) panel shows the flux ratio in the energy range in which 99\% of the DeepCore (IceCube) events are contained.
   }\label{fig:flux_ratio}
\end{figure}%TODO: make prettier cmap

@@

The data include 14901 track-like events and 26001 cascade-like events, both divided into eight 
$ \log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins. Each event has a Monte Carlo weight $w_{ijk,\beta}$
from which we can construct the event count as
\begin{align}\label{eq:MCevents}
    N_{ijk} &= C_{ijk}\sum_{\beta}w_{ijk,\beta}\, \phi_\beta^\text{det}\,,
\end{align}
where $C_{ijk}$ is the correction factor from the detector systematic uncertainty and $\phi_\beta^\text{det}$ is defined as Eq.~\ref{eq:propFlux}. We have now substituted the effect of the Gaussian smearing 
by treating the reconstructed and true quantities as a migration matrix.

@@

We treat the uncorrelated systematic uncertanties differently for each detector. For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
We consider best, normal, and worst-case scenarios in IceCube using
$f=5\%$, $10\%$, and $15\%$ respectively. For PINGU, we use the same form but instead use $f=0\%$, $3\%$, and $5\%$. %TODO: maybe source/justification for these values?
For DeepCore, we use the provided systematic error distribution which accounts for uncertainties in the finite MC statistics and the data-driven 
muon background estimate~\cite{DC2019data}. This is summarized in Table~\ref{table:syst_errors}.  %TODO segue
{\renewcommand{\arraystretch}{1.2}
\begin{table}
   \begin{tabular}{lccc}
      \hline \hline
      Experiment & Best case & Baseline & Worst case \\
      \hline
      IceCube & $5\%$ & $10\%$ & $15\%$ \\
      PINGU & $0\%$ & $3\%$ & $5\%$ \\
      \hline \hline
   \end{tabular}
   \caption{Definition of the best, baseline, and worst case scenarios considered in each experiment, modelled by $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$ with $f$ from the table.
   We do not consider different DeepCore scenarios because her systematic error distribution is already provided in the data release~\cite{DC2019data}.}\label{table:syst_errors}
\end{table}

\subsection{Constraining the NSI parameters}
First, we set all standard oscillation parameters to their current best-fit values of Eq.~\ref{eq:PINGUparams}, except for $\dm$ and $\theta_{23}$, 
which we marginalize over their $3\sigma$ ranges of \SI{2.435e-3} to \SI{2.598e-3}{\electronvolt^2} and \SI{40.1} to \SI{51.7}{\degree} respectively. %TODO: make unit ranges nicer.
After the oscillation parameters have been marginalized out, we plot $\Delta \chi^2$ for each of the four NSI parameters in Fig.~\ref{fig:3D_NO}. We record the 
confidence levels in Table~\ref{table:IC_DC_results} and Table~\ref{table:PINGU_joint_results}, and best-fit points in Table~\ref{table:bestfit}.

@@

We see asymmetrical shapes in the $\Delta \chi^2$ plots in Fig.~\ref{fig:3D_NO}, especially in $\emt$ and $\eet$.%TODO: make better
We investigate this for $\emt$ by plotting the muon survival probability difference $P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$, where 
%TODO: check that this is what is plotted.
$\epsilon^\pm_{\mu\tau} = \pm 0.01$ in Fig.~\ref{fig:Pmm_asymmetry}.
Muon events are the most abundant, and it suffices to study $P_{\mu\mu}$ to explain the $\emt$ asymmetry. 
Let $P^\pm$ denote $P_{\mu\mu}(\epsilon^\pm_{\mu\tau})$. 
We see see that $\emt^+$ generates a slightly higher $P_{\mu\mu}$ for energies around \SI{20}{\GeV} (blue area),
while $\emt^-$ produces higher $P_{\mu\mu}$ for almost all other combinations of ${\Etrue,\ztrue}$. As we see in Fig.~\ref{fig:flux_ratio}, this muon survival abundance 
is indeed preserved at flux level. So the flux for $\emt^+$ is higher than the flux for $\emt^-$. Is this still true at event level, i.e. after reconstruction? 
As we see in Fig.~\ref{fig:emt_events}, the binned PINGU event counts display strong differences for many bins, giving a high $\Delta \chi^2$ on both sides of $\emt=0$ and slightly higher for $\emt^-$. DeepCore on the other hand, 
has fewer bins where the event count for $\emt^-$ surpasses the event count for $\emt^+$, giving weaker statistics for the negative side. Thus, we conclude that the DeepCore
$\emt$ asymmetry stems from the lower statistics for negative $\emt$ at probability level, which is then propagated through the flux and finally affects the reconstruction. %TODO: fill more conclusion?

\begin{figure}
   \begin{subfigure}{0.25\textwidth}   
      \includegraphics[width=1\linewidth]{figures/Pmm_asymmetry.pdf}%TODO: redo for emt=0.01x
      \caption{$P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$ using  $\epsilon^\pm_{\mu\tau} = \pm 0.01$}\label{fig:Pmm_asymmetry}
   \end{subfigure}
   \begin{subfigure}{0.6\textwidth}
      \includegraphics[width=1\linewidth]{figures/emt_events.pdf}
   \caption{The best-fit event count ratio from each detector for $\emt = \pm 0.01$. IceCube only observes a small surplus of events for $\emt=-0.01$ compared to 
   DeepCore and PINGU due to the weak NSI effect at high \si{\GeV} energies.}\label{fig:emt_events}%TODO: improve caption?
   \end{subfigure}
   \caption{Probability difference and best-fit event count. Negative $\emt$ has the strongest signal, as the thin blue 
   region stemming from $\emt = 0.01$ is less prominent than the red regions from $\emt=-0.01$.}
\end{figure}

@@

For the joint analysis, we follow the parameter goodness-of-fit prescription~\cite{maltoni2003} and construct the joint $\chi^2$ as 
\begin{align}\label{eq:joint_chisq}
    \chi^2_\text{joint} = \sum_\text{exp}\chi^2_\text{exp} - \chi^2_\text{exp,min}\,
\end{align}
with test statistic $\chi^2_\text{joint,min}$. The $\Delta \chi^2_\text{joint}$ is then $\Delta \chi^2_\text{joint} = \chi^2_\text{joint} - \chi^2_\text{joint,min}$.
The results are shown in Fig.~\ref{fig:IC_NO} and summarized in Tables~\ref{table:IC_DC_results} and~\ref{table:PINGU_joint_results}.
Comparing the PINGU and the DeepCore results in Fig.~\ref{fig:3D_NO}, we note that the best-fit for each NSI parameter for the PINGU experiment is expected to be zero. This is because the `data' we generated during 
the PINGU simulations assume no NSI since they have yet to be observed in nature. This introduces a non-NSI bias in all joint analyses which include PINGU,
since PINGU has stronger statistics than DeepCore and will thus pull the joint $\chi^2$ towards $\epsilon =0$.
\newpage
